---
title: "Exercise 7 Solution"
author: "Cameron Bale"
format: docx
---

## Exercise 7 | Regional Spending Differences

The retailer asks: **Do customers in different regions spend differently, and do those differences persist after accounting for household characteristics?**

Use `home_goods_customers.csv` and `data_dictionary.pdf`. Your outcome is `PreviousSpend`, spending over the previous 12 months. Choose and report a baseline for `Region` and any other categorical predictor.

Compare **two to three models** that account for regions and some demographic characteristics of your choosing.

Use the same customers for each comparison and report the sample size. Use clear units, such as thousands of dollars for income. Choose the code operations and their sequence.

## Exercise 7 | Evidence for the Client {.smaller}

Submit a Word report rendered from Quarto with your code, results and interpretation. Include:

- Your models, reasons for the household variables you selected, and each model's $R^2$.
- A coefficient table with 95% intervals and a plot of the point estimates and confidence intervals of your preferred model. Interpret one regional comparison and one continuous slope, holding other included variables fixed.
- A comparison of the regional estimates before and after accounting for household characteristics. Explain what changes, how omitted variables could affect the comparison, and what the intervals tell you.
- A check for multicollinearity using VIF after fitting your models. Explain any concerns and justify one decision to include or exclude a candidate variable.

## Exercise 7 | Compare the Same Household Across Regions

Use your preferred model to predict spending in **all four regions** for a household with **characteristic values of your choosing**.

Choose and report values for the characteristics included in your model. Keep those values fixed across regions, changing only `Region`. Explain the predicted regional differences alongside their coefficient uncertainty.

Recommend whether region is useful for describing customer value. Identify one plausible influence still omitted, and explain why the regional comparisons do not establish what a change in location or an email campaign would cause.

Upload your Word report to Canvas's **Exercise 7** assignment.

## Grading Rubric | 5 Points

Award one point for each row; partial credit is appropriate for a substantially correct analysis with a small error. Different baselines, household predictors and defensible model choices can receive full credit.

| Point | Full-credit evidence |
|---|---|
| 1. Model choices and fit | Two to three models address the regional question using demographic characteristics of the student's choosing, with reasons for their selections. Categorical baselines, units, common comparison sample and sample size are reported; R-squared values are correctly explained. |
| 2. Coefficients and uncertainty | The coefficient table with 95% intervals and a plot of the point estimates and confidence intervals of the preferred model are correct. One regional comparison and one continuous slope are interpreted in dollars, holding other included predictors fixed. |
| 3. Regional comparison and omitted variables | Regional coefficients before and after adjustment are compared. The explanation distinguishes the two comparisons, considers an omitted-variable story and uses intervals without claiming that any coefficient change proves causality. A small change or uncertain regional difference is a valid finding. |
| 4. Predictor overlap and judgment | VIFs are correctly assessed after fitting the models, and one inclusion or exclusion decision is justified using the business story and evidence. Full credit does not require a high VIF or dropping a variable. |
| 5. Profiles and recommendation | The student chooses and reports values for the characteristics in their model, holds those values fixed across all four regions, and computes correct predictions. The Word report gives an evidence-based recommendation, a plausible remaining omitted influence and the distinction between associations and intervention effects. |

## Example Solution | A Regional Customer Story

This example uses **West** as the regional baseline. It accounts for income, household size, homeownership and home size: four characteristics representing household resources and demand for home goods. A third model checks whether website activity adds useful information. Other justified choices are acceptable.

```{r}
library(tidyverse)
library(tidymodels)

customers <- read_csv("home_goods_customers.csv", show_col_types = FALSE)

region_data <- customers |>
  drop_na(PreviousSpend, Region, HouseholdIncome, HouseholdSize,
          Homeowner, HomeSizeSqFt, SiteVisits30d, ProductViews30d) |>
  mutate(
    IncomeThousands = HouseholdIncome / 1000,
    HomeSizeHundreds = HomeSizeSqFt / 100,
    RegionNortheast = if_else(Region == "Northeast", 1, 0),
    RegionMidwest = if_else(Region == "Midwest", 1, 0),
    RegionSouth = if_else(Region == "South", 1, 0)
  )

nrow(region_data)
```

All three models use **9,603 customers**. West customers have zero for all three regional indicators. The other regional coefficients compare their region with West. Using an intercept and all four indicators would be redundant.

## Compare Regions, Then Account for Household Characteristics

```{r}
fit_region <- linear_reg() |>
  set_engine("lm") |>
  fit(PreviousSpend ~ RegionNortheast + RegionMidwest + RegionSouth,
      data = region_data)

fit_household <- linear_reg() |>
  set_engine("lm") |>
  fit(PreviousSpend ~ RegionNortheast + RegionMidwest + RegionSouth +
        IncomeThousands + HouseholdSize + Homeowner + HomeSizeHundreds,
      data = region_data)

fit_candidate <- linear_reg() |>
  set_engine("lm") |>
  fit(PreviousSpend ~ RegionNortheast + RegionMidwest + RegionSouth +
        IncomeThousands + HouseholdSize + Homeowner + HomeSizeHundreds +
        SiteVisits30d + ProductViews30d,
      data = region_data)

model_fit <- bind_rows(
  region = glance(fit_region), household = glance(fit_household),
  browsing = glance(fit_candidate), .id = "model"
) |>
  select(model, r.squared)
model_fit
```

Region alone explains about **0.02%** of spending variation. The household-adjusted model explains about **51.2%**; adding browsing leaves it at approximately **51.2%**. Household information is much more useful here than region alone. Higher R-squared by itself does not establish a causal relationship or better predictions for new customers.

## What Changes in the Regional Comparison?

```{r}
regional_comparison <- bind_rows(
  region = tidy(fit_region, conf.int = TRUE),
  household = tidy(fit_household, conf.int = TRUE), .id = "model"
) |>
  filter(term %in% c("RegionNortheast", "RegionMidwest", "RegionSouth")) |>
  select(model, term, estimate, conf.low, conf.high)
regional_comparison
```

Relative to West, the initial differences are about **−$1.98 for Northeast**, **−$2.51 for Midwest**, and **$4.73 for South**. After accounting for the four household characteristics, they are about **$5.27**, **−$3.69**, and **$3.03**, respectively. All of these regional intervals include zero.

The first model compares the customers living in each region, including differences in their household profiles. The second compares regions while holding the included household characteristics fixed. For example, income could affect spending and differ across regions; omitting it could mix household resources into a regional coefficient. Here, the regional differences are small and uncertain both before and after adjustment. We should not claim a strong regional relationship or dramatic omitted-variable bias from this comparison.

## Interpret the Preferred Model

The household-adjusted model is preferred in this example: it answers the regional question, describes household spending differences and avoids adding two overlapping browsing measures that barely improve fit.

```{r}
household_estimates <- tidy(fit_household, conf.int = TRUE)
household_estimates

household_estimates |>
  filter(term != "(Intercept)") |>
  ggplot(aes(x = term, y = estimate)) +
  geom_point() +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = .1) +
  geom_hline(yintercept = 0, color = "red") +
  coord_flip()
```

Holding the other included characteristics fixed, Northeast customers spend about **$5.27 more than West customers**, with a 95% interval of approximately **−$4.41 to $14.95**. The estimate does not provide clear evidence of a difference. Another **$1,000 of annual income** is associated with about **$2.29 more spending**, with an interval of approximately **$2.23 to $2.35**.

The intercept describes West customers with zero income, zero household size, zero home size and no homeownership. That is not a realistic customer profile. The confidence intervals quantify uncertainty about coefficients, not the range of individual spending.

## Check Overlap and Make a Variable Decision

```{r}
fit_candidate |>
  extract_fit_engine() |>
  car::vif() |>
  enframe(name = "predictor", value = "VIF") |>
  arrange(desc(VIF))

fit_household |>
  extract_fit_engine() |>
  car::vif() |>
  enframe(name = "predictor", value = "VIF") |>
  arrange(desc(VIF))
```

The `car` package must be installed to use `car::vif()`, as in the lecture.

Visits and views have VIFs of about **6.20** and **5.81** in the candidate model, exceeding the lecture's guideline of 4. They contain strongly overlapping information, making their separate slopes harder to estimate precisely. All VIFs in the preferred household model are below **1.9**.

This example excludes both browsing measures: neither is needed for the household-comparison question, they barely change fit, and retaining both adds an interpretation problem.

## Predict the Same Household Across Regions

For this example, choose annual income of $80,000, household size of three, homeownership and home size of 1,600 square feet. These are values for the characteristics included in the preferred model. Students may choose other values; hold the chosen profile fixed across regions.

```{r}
profiles <- tibble(
  Region = c("West", "Northeast", "Midwest", "South"),
  RegionNortheast = c(0, 1, 0, 0),
  RegionMidwest = c(0, 0, 1, 0),
  RegionSouth = c(0, 0, 0, 1),
  IncomeThousands = 80,
  HouseholdSize = 3,
  Homeowner = 1,
  HomeSizeHundreds = 16
)

profile_predictions <- bind_cols(
  profiles, predict(fit_household, new_data = profiles)
) |>
  select(Region, .pred)
profile_predictions
```

Predicted spending is roughly **$1,000 in West**, **$1,005 in Northeast**, **$997 in Midwest**, and **$1,003 in South**. The household inputs are identical, so differences equal the corresponding regional coefficients. These small point differences should be read alongside the intervals; they do not establish that one region has a higher expected value with confidence.

## Recommendation and Remaining Questions

These results give little support for prioritizing customers simply because of their region. Household characteristics explain substantially more spending variation. Region may still matter for other decisions, such as shipping costs, which this spending model does not address.

**Local store access** is one plausible omitted influence: access may vary across regions and affect how easily customers shop with this retailer. The broad region labels and included household characteristics do not capture it. Other well-explained omitted influences are acceptable.

The analysis describes associations in historical spending. It does not establish what moving the same household to another region would cause, or how that household would respond to an email campaign.
